---
title: "TDP43 concentration dependent analysis"
author: "Matteo Zanovello"
date: "2022-01-10"
output:
---

```{r setup, include = False}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
if (!require("pacman")) {install.packages("pacman")}
if (!requireNamespace("BiocManager", quietly = TRUE)) {install.packages("BiocManager")}
pacman::p_load(tidyverse, data.table, ggplot2, janitor, ggpubr, ggrepel,
               devtools, readxl, DESeq2, readr, edgeR, annotables,
               PCAtools, clusterProfiler, "org.Hs.eg.db", enrichplot, pcaExplorer, topGO, UpSetR)

source(here::here('scripts','create_feature_count_table.R'))
source(here::here('scripts','make_deseq_dfs.R'))
source(here::here('scripts','run_standard_deseq.R'))
source(here::here('scripts','make_volcano_plot.R'))
source(here::here('scripts','splicing_volcanoplot.R'))
source(here::here('scripts','run_de_irfinder.R'))
```

this organizes all the data from the many .bam files and creates one data frame you can work with,
then runs the make_deseq_dfs() function and run_standard_deseq() and eventually make graphs
```{r}
featureCounts <- create_feature_count_table("data/feature_count") #change here!!

grepper <- c("a", 
             "b", 
             "c", 
             "f", 
             "h"
             )
names(grepper) <- c("dox00125", 
                    "dox00187", 
                    "dox0021", 
                    "dox0025", 
                    "dox0075"
                    )
for (j in grepper) {
my.dds <- run_standard_deseq("data/feature_count", #change here!!
                             base_grep = "NT",
                             contrast_grep = "DOX",  
                             grep_pattern = j,
                             baseName = "Untreated",
                             contrastName = 'TDP43KD')

bbb <- label_significant(my.dds$results_table, log2FoldCut = 2, log10padj = 10) + annotate("text", x=4, y=-0.25, label=names(j))
plot(bbb)
}
```

splicing volcano
```{r}
base <- "noDox"
contrast <- c("dox00125", 
              "dox00187", 
              "dox0021", 
              "dox0025", 
              "dox0075"
              )
for (i in contrast) {
    input <- paste(base, i, sep = "-")
    input_csv <- paste(input, "csv", sep = ".")
input_splicing <- fread(file.path(here::here(), "data/psi", input_csv))
input_splicing <- separate_rows(input_splicing, c(names(input_splicing[,c(4:9,12,15)])), sep  = c(";"), convert = T)

a <- splicing_dots_tables_function(input_splicing) #+ annotate("text", x=0.9, y=0.1, label=i)
plot(a)
}
```

This is the vignette for the PCAtools package
https://bioconductor.org/packages/release/bioc/vignettes/PCAtools/inst/doc/PCAtools.html

We're going to put the "ensgene" into a 'rowname' and then do the PCA analysis using meta data
Now I'm going to write this using the pipe operator "%>%". Your "hot key" to type that in is: Ctrl + Shift + M
```{r}
colnames(featureCounts) <- c("ensgene", 
                           "DOX_0.0125_1", "DOX_0.0125_2", "DOX_0.0125_3",
                           "DOX_0.0187_1", "DOX_0.0187_2", "DOX_0.0187_3",
                           "DOX_0.021_1", "DOX_0.021_2", "DOX_0.021_3", 
                           "DOX_0.025_1", "DOX_0.025_2", "DOX_0.025_3", 
                           "DOX_0.075_1", "DOX_0.075_2", "DOX_0.075_3",
                           "NT_0_1", "NT_0_2", "NT_0_3", 
                           "gene_name")

pca_table <- featureCounts[,1:19] %>%
    column_to_rownames('ensgene')
    
path_to_meta_you_downloaded <-  "data/metadata.csv" #update this
meta_df <- fread(path_to_meta_you_downloaded)
meta_df <- meta_df %>%
    column_to_rownames('sample') 
meta_df[,1] <- as.factor(meta_df[,1])
meta_df[,2] <- as.factor(meta_df[,2])

annotation <- featureCounts[,c(1,20)] %>%
    column_to_rownames('ensgene')

pcaExplorer(countmatrix = pca_table, coldata = meta_df, annotation = annotation)

```

after running IRFinder, compute DE
```{r}
ir_top_level = "data/irfinder/"
experiment = make_deframe_irfinder(ir_top_level,base_grep = "NT",
                                   contrast_grep = "DOX")$metadata
paths = make_deframe_irfinder(ir_top_level,base_grep = "NT",
                                   contrast_grep = "DOX")$path

for (j in c("0.0125", "0.0187", "0.021", "0.025", "0.075")) {
    experiment <- filter(experiment, grepl(j, SampleNames) | Condition == "control")
    paths <- paths[grepl(j, paths) | grepl("NT", paths)]
}

dds = run_deseq_ir(paths,experiment = experiment)
result = return_formated_results(dds)
result0075 <- result %>% 
    mutate(ir_change = TDPKD -  control)

result_merged <- dplyr::bind_rows(list(result00125, result00187, result0021, result0025, result0075), .id = 'source')
```



GO analysis https://yulab-smu.top/biomedical-knowledge-mining-book/clusterprofiler-go.html
#see notes above
```{r}
results <- my.dds$results_table
filtered_res <- filter(results, padj < 0.1 & abs(log2FoldChange) > 2)
## feature 1: numeric vector
geneList <- pull(filtered_res,3)
## feature 2: named vector
names(geneList) = as.character(pull(filtered_res,8))
## feature 3: decreasing order
geneList = sort(geneList, decreasing = TRUE)

ggo <- groupGO(gene = names(geneList),
               OrgDb = org.Hs.eg.db,
               keyType = "SYMBOL",
               ont = "MF",
               level = 3)
cnetplot(ggo)

ego <- enrichGO(gene = names(geneList),
                keyType = "SYMBOL",
                OrgDb = org.Hs.eg.db,
                ont = "BP",
                pAdjustMethod = "BH",
                pvalueCutoff = 0.01,
                qvalueCutoff = 0.05)
goplot(ego)

pgo <- pairwise_termsim(ego)
emapplot(pgo)

gse <- gseGO(geneList     = geneList,
              keyType = "SYMBOL",
              OrgDb        = org.Hs.eg.db,
              ont          = "CC",
              minGSSize    = 50,
              maxGSSize    = 200,
              pvalueCutoff = 0.01,
              verbose      = FALSE)
```